# Install
```{r}
## One-time install
install.packages("dartR")
install.packages("BiocManager")
install.packages("devtools")
BiocManager::install(c("SNPRelate", "qvalue"))
gl.install.vanilla.dartR()
```

# Import and check genlight object
```{r}
# import data
library(dartR)

gl <- gl.read.dart(
  filename = "Report_DPoc22-7677_SNP_2.csv",
  ind.metafile = "ind_metrics.csv" # csv file which contains metrics associated with each individual (id, pop, sex, etc.)
)

# filter out individuals we are not using
# gl <- gl.drop.ind(gl, ind.list = c("89", "146", "139", "140", "102", "142"))

# to check genlight object is compliant
gl <- gl.compliance.check(gl)

# save(gl, file="gl_dartR.rdata")
```

```{r}
# some useful commands to get data from your genlight object

nLoc(gl) # number of loci

locNames(gl) # list of loci

nInd(gl) # number of individuals (specimens / samples)

indNames(gl) # list of individuals

nPop(gl) # number of populations

popNames(gl) # list of populations

pop(gl) # list of population assignments for each individual

as.matrix(gl) # generate a matrix of the SNP scores, with 0 as homozygous reference, 2 as homozygous alternate, and 1 as heterozygous.

glPlot(gl) # a smear plot of individual against locus, useful for gross pattern identification and assessment of allelic dropout
```

# Filtering data
```{r}
## 1. FILTER SECONDARIES
# gl.report.secondaries(gl_f) # report
gl_f <- gl.filter.secondaries(gl, verbose = 3) # this is standard, no threshold set

## 2. FILTER REPRODUCIBILITY
# gl.report.reproducibility(gl_f) # report
gl_f <- gl.filter.reproducibility(gl_f, threshold = 0.99, verbose = 3) # filtering threshold 0.99

## 3. FILTER CALL RATE BY LOCI
# gl.report.callrate(gl_f, method = "loc") # report
gl_f <- gl.filter.callrate(gl_f, method = "loc", threshold = 0.60, verbose = 3) # filtering
# Note: this has been changed to 0.6 due to significant loss of data at higher thresholds

## 4. FILTER CALL RATE BY INDIVIDUALS
# gl.report.callrate(gl_f, method = "ind") # report
gl_f <- gl.filter.callrate(gl_f, method = "ind", threshold = 0.80, verbose = 3) # filtering

## 5. FILTER READ DEPTH TO REMOVE LOW COVERAGE LOCI
# gl.report.rdepth(gl_f) # report
gl_f <- gl.filter.rdepth(gl_f, lower = 5, upper = 85, verbose = 3) # filtering
# selecting a threshold of 10% either side (data between the 10th and 90th percentile was kept)

## 6. FILTER MONOMORPHIC LOCI
# gl.report.monomorphs(gl_f) # report
gl_f <- gl.filter.monomorphs(gl_f, verbose = 3) # filtering
# Filtered prior to MAF as it was generating a warning about monomorphoc loci

## 7. FILTER Minor Allele Frequencies (MAF)
# gl.report.maf(gl_f) # report
gl_f <- gl.filter.maf(gl_f, threshold = 0.01, verbose = 3) # filtering

# filtering complete.

gl_f <- gl.recalc.metrics(gl_f) # re-calculate locus metadata now that filtering is complete
```


# Test for clones - none under the threshold 0.45
```{r}
# 80% & 70% loci coverage
gl_grm <- gl.grm(gl_f, verbose = 3)

# This function calculates the mean probability of identity by state (IBS) across loci that would result from all the possible crosses of the individuals analyzed.
# IBD is calculated by an additive relationship matrix approach developed by Endelman and Jannink (2012)

# Represents a genomic relationship matrix (GRM) as a network
gl_grm_net <- gl.grm.network(gl_grm, gl_f, method = "mds", link.size = 2, relatedness_factor = 0.45) # decided on value between sibling/half sibling
```

# Impute missing data using the nearest neighbour function - needed for some downstream analyses (e.g. Fst)
```{r}
gl_f2 <- gl.impute(gl_f, method = "neighbour", verbose = 5)
# re-calculating the locus metrics after manipulation
gl_f2 <- gl.recalc.metrics(gl_f2)

# 2244 / (nInd(gl_f) * nLoc(gl_f)) # Percentage imputed
```

# Check for Symbiodiniaceae contamination - no symbiont sequences aligned with filtered data indicating no contamination
```{r}
# Download Symbiodiniaceae genomes here: sampgr.org.cn/index.php/download
# Download & install BLAST (needed to run analysis) https://ftp.ncbi.nlm.nih.gov/blast/executables/blast+/LATEST/
# select this file: ncbi-blast-2.14.1+.dmg
# Run the BLAST function for each genome
gl.blast(gl_f, "Symbiontgenomes/symC_scaffold_40.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/Symbiodinium_microadriacticum_genome.scaffold.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/symA3_scaffold_37.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/Fugacium_kawagutii_V3_genome_scaffold.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/Fugacium_kawagutii_V2_genome_Scaffolds.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/Fugacium_kawagutii_V1_genome_scaffold.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/Cladocopium_goreaui_Genome.Scaffolds.fasta", verbose = 5)
gl.blast(gl_f, "Symbiontgenomes/Breviolum_minutum.v1.0.genome.fa", verbose = 5)
```


#PCA
```{r}
library(adegenet)
source("../default_theme.r")

pca_data <- gl_f

# levels(pca_data$pop) = c(
#   "L1 Flat", "L1 Pale" ,"L2 Slope", "Low Isles slope pale",
#   "L2 Flat", "O2 Crest", "O2 Slope", "O3 Flat", "O3 Slope" ,
#   "T1 Slope", "T2 Slope", "O1 Slope"
# )

pca_data <- gl.rename.pop(pca_data, old = "Low Isles flat", new = "L1 Flat")
pca_data <- gl.rename.pop(pca_data, old = "Low Isles flat pale", new = "L1 Pale")
pca_data <- gl.rename.pop(pca_data, old = "Low Isles Woody flat", new = "L2 Flat")
pca_data <- gl.rename.pop(pca_data, old = "Low Isles slope", new = "L2 Slope")
pca_data <- gl.rename.pop(pca_data, old = "Tongue Turtle Bay", new = "T2 Slope")
pca_data <- gl.rename.pop(pca_data, old = "Tongue Blue Hole", new = "T1 Slope")
pca_data <- gl.rename.pop(pca_data, old = "Opal Long Bommie", new = "O1 Slope")
pca_data <- gl.rename.pop(pca_data, old = "Opal Mojo flat", new = "O2 Crest")
pca_data <- gl.rename.pop(pca_data, old = "Opal Mojo slope", new = "O2 Slope")
pca_data <- gl.rename.pop(pca_data, old = "Opal Rayban flat", new = "O3 Flat")
pca_data <- gl.rename.pop(pca_data, old = "Opal Rayban slope", new = "O3 Slope")

pcoa <- gl.pcoa(pca_data, verbose = 3)

# gl.pcoa.plot won't return the plot for us to work with
# We can tell it to save it in a temporary directory though
gl.pcoa.plot(pcoa, pca_data, pop.labels = "pop", ellipse = FALSE, save2tmp = TRUE)
# Now we need to get the list of repots it has saved and find the last plot
gl.list.reports()
# put the number of that plot here so we can finally get the plot and work with it
plot <- gl.print.reports(5)

# Remove the horizontal and vertical lines
plot$layers[[4]] <- NULL
plot$layers[[3]] <- NULL

# output the plot
plot +
  # stat_ellipse(type = "norm", level = 0.95) +
  xlim(-5, 10) + ylim(-10, 10) +
  default_theme

ggsave("Genetic Analysis/plots/pca.png", width = 6, height = 5)
# ggsave("Genetic Analysis/plots/pca_longbommie.png", width = 6, height = 5)


###### To find out which samples are outliers ###### - this only needs to be run once

# Similar to how we get the plot, the results must be found by running gl.list.reports() and finding the latest Table number
pca_results <- gl.print.reports(16)

# Load the metadata so we can get Sample IDs
metadata <- read_csv("Genetic Analysis/ind_metrics_site.csv") %>%
  # Create a `ind` column to match the PCA output
  mutate(ind = as.character(id))


outliers <- pca_results %>%
  # Join the metadata so we can get Sample IDs
  left_join(metadata, "ind") %>%
  # We can see from the plot there are some outliers above PC2 = 4 and below PC2 = -4
  filter(PCoAy > 4 | PCoAy < -4)
```

# STRUCTURE analysis 
```{r}
library(stringr) # for str_trunc

# gl.run.structure takes a genlight object and runs structureRun (https://rdrr.io/cran/dartR/man/gl.run.structure.html)
#     ↓↓
# structureRun expects a gtypes object in R and runs STRUCTURE from inside R (https://rdrr.io/cran/strataG/man/structure.html)
#     ↓↓
# STRUCTURE is a program external to R
# The program: ./STRUCTURE/structure
# Documentation: https://web.stanford.edu/group/pritchardlab/structure_software/release_versions/v2.3.4/structure_doc.pdf

# For this function to work, make sure that individual and population names have no spaces.
# make a copy od the data to make changes to
gl_for_structure <- gl_f

# gl.run.structure below works but R has no way to capture stdout from it and
# there's a lot of output so it overflows the terminal and we can't ever get it
# The solution is to output gl_for_structure to file, and run gl.run.structure
# in another R session piping the output to a file
save(gl_for_structure, file = "gl_for_structure.RData")
# now run `Rscript structure_run.r > outputs/structure_output.txt` in terminal
# Now we load the results of gl.run.structure from structure_run.r
load("structure_run.RData") # loads a variable called sr

# This is the code that's now run inside structure_run.r
# sr <- gl.run.structure(
#   gl_for_structure,
#   burnin = 10000,
#   numreps = 10000,
#   save2tmp = TRUE,

#   # Documentation of options here: https://rdrr.io/cran/strataG/man/structure.html#heading-2
#   k.range = 1:10,
#   num.k.rep = 5, # number of replicates for each value in k.range.

#   # The location of the external STRUCUTRE program to run
#   exec = "./STRUCTURE/structure"
# )

# Calculate Evanno metrics
ev <- gl.evanno(sr)
ev
# Creates Evanno plots
evanno_plots <- gl.evanno(sr)

source("../.r")

# delta.k
evanno_plots$plots$delta.k + default_theme
ggsave("Genetic Analysis/plots/evanno_delta_k.png")

# ln.pk
evanno_plots$plots$ln.pk + default_theme
ggsave("Genetic Analysis/plots/evanno_ln_pk.png")

# ln.ppk
evanno_plots$plots$ln.ppk + default_theme
ggsave("Genetic Analysis/plots/evanno_ln_ppk.png")

# mean.ln.k
evanno_plots$plots$mean.ln.k + default_theme
ggsave("Genetic Analysis/plots/evanno_mean_ln_k.png")
```

# AMOVA
```{r}
# gl.amova(gl_f)

library(poppr)

# Converting genlight to genind
gen_ind <- gl2gi(gl_f)
# Assign pop information to strata
strata(gen_ind) <- data.frame(pop = pop(gen_ind))
# Performing AMOVA
amova <- poppr.amova(gen_ind, ~pop, nperm = 1000, within = TRUE)
amova

# Test for significance
amova_sig <- randtest(amova, nrepet = 999)
plot(amova_sig)
amova_sig
```

# ADMIXUTRE analysis
```{r}
# Converting genlight to PLINK format
gl2plink(gl_f, plink_path = "./PLINK", bed_file = TRUE, verbose = 5, outpath = "./PLINK")
# The file "PLINK/gl_plink.bed" has been created and exported out of R for use in shell
# Note PLINK folder moved up one level in folders as it does not like spaces in file names

# create ind2pop.txt needed for pong
ind2pop <- pop(gl_f) # Get population names for each individual
levels(ind2pop) <- gsub(" ", "", levels(ind2pop)) # Remove spaces from factor levels
write.table(ind2pop, "Genetic Analysis/outputs/ind2pop.txt", row.names = FALSE, col.names = FALSE, quote = FALSE)
# The file "ind2pop.txt" has been created and exported out of R for use in shell
# Saved to outputs folder in Genetic Analysis - not saved in PLINK folder

# In terminal run admixture using the exported files from above - https://speciationgenomics.github.io/ADMIXTURE/
# Use ADMIXTURE/ADMIXTURE.sh to run ADMIXTURE and pong (plots)
```


# Genetic distance matrix between populations
```{r}
# Genetic distance between two individuals or populations is a measure of their genetic dissimilarity. large value = very dissimilar, small value = similar, 0 = identical genetic profiles
# gl.dist.ind() or gl.dist.pop()
# Note: Method = “euclidean” is severely affected by missing values and should be used only for complete data. Therefore, use "gl_f2" which has filled in any missing data via the nearest neighbour function

Dist_pop <- gl.dist.pop(gl_f2, method = "euclidean")
Dist_pop

# Heat map
svg("Genetic Analysis/plots/heatmap.svg")
gl.plot.heatmap(Dist_pop,
  key = TRUE, keysize = 1.5, key.title = NA, key.ylab = NA,
  cexRow = 1, cexCol = 1, margins = c(10, 10), srtCol = 45
)
dev.off()

# Calculate Fst values
# Fst is a measure of population differentiation due to genetic structure. An Fst value greater than 0.15 can be considered as significant in differentiating populations (Frankham et al., 2002).
Fst <- gl.fst.pop(gl_f2, nboots = 999) # note: don't need to include nboots if I'm not interested in p-values
Fst

write.csv(Fst$Fsts, "Genetic Analysis/outputs/Fst.csv")
write.csv(Fst$Pvalues, "Genetic Analysis/outputs/FstPvalues.csv")
```

### Fixed difference analysis
      
```{r}
D <- gl.fixed.diff(gl_f, v = 4)
# Warning: Fixed differences can arise through sampling error if sampling sizes are small

D$fd
# Note that there are a few comparisons (IM-OM and IR-OR) with zero fixed differences. These are candidates to aggregate on the basis of presenting no diagnostic allelic differences.

# Testing for Significance
test_results <- gl.fixed.diff(D, test = TRUE, v = 3)

test_results$pval

# NOTE
# Note continuing from here as there are only 2 populations

# Aggregate

# At this point we might consider aggregating sample sites pairwise where they have not accumulated any fixed differences.

D2 <- gl.collapse(D, tpop = 1, verbose = 3)

# Population with 0 fixed differences are now grouped together. Inner and outer mangrove in Inner mangrove+ and Inner and outer reef in Inner reef+



D2$fd

# This is as far as the collapsing of the fixed difference matrix can go, with tpop set to 1.



# Testing for Significance: there is one last issue to consider, the possibility that distinctions between our final aggregations are based on false positives (given the sample/finite sample size). Note that my populations have sample sizes of 22 (mangrove) and 18 (reef).

D4 <- gl.fixed.diff(D2, test = TRUE, v = 3)

D4

# Outcome: Inner mangrove+ vs Inner reef+  [p = 1 ,ns]

# The 2 fixed allelic differences between Inner mangrove+ and Inner reef+ don't exceed the false positive rate, probably given the exceptionally low sample sizes (despite the fact that we grouped them).
```


# Run 'popgen' (snpReady) for genetic characterisation
```{r}
library(snpReady)

# Get current populations from gl_f
populations <- as.matrix(pop(gl_f))

# Run popgen with subgroups equal to my 'populations'
popgen_output <- popgen(
  as.matrix(gl_f),
  subgroups = populations
)

means <- NULL

for (pop in names(popgen_output$bygroup)) {
  pop_data <- popgen_output$bygroup[[pop]]
  print(pop) # prints name of the 'population'
  # print(pop_data$Genotypes)
  # print(pop_data$Markers)
  # print(pop_data$Population)
  # print(pop_data$Variability)

  # calculate metrics of interest - we are using genetic diversity (GD)
  means <- rbind(means, data.frame(
    pop = pop,
    p = mean(pop_data$Markers$p), # allele frequency (frequency of dominant allele - p)
    q = mean(pop_data$Markers$q), # allele frequency (frequency of recessive allele - q)
    MAF = mean(pop_data$Markers$MAF), # minor allele frequency
    He = mean(pop_data$Markers$He), # observed heterozygosity
    Ho = mean(pop_data$Markers$Ho), # expected heterozygosity
    GD = mean(pop_data$Markers$GD), # Nei's genetic diversity
    PIC = mean(pop_data$Markers$PIC), # polymorphic information count
    Miss = mean(pop_data$Markers$Miss), # missing rate,
    n = nrow(pop_data$Genotypes) # gives sample number for each group
  ))
}

print(means)
write.csv(means, "Genetic Analysis/outputs/GeneticDiversity.csv")
```

